---
title: "classification"
author: "Wayne Lee"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Please download the "classification.Rmd" from CourseWorks + the music*.csv file!
code_from_lecture/classification.Rmd

## Music dataset

From the [CORGIS data project](https://corgis-edu.github.io/corgis/) I've obtained a [music dataset](https://corgis-edu.github.io/corgis/csv/music/).

```{r}
music <- read.csv("~/Downloads/music_hottest_song_only.csv")
y <- as.numeric(music$artist.terms == "hip hop")
x <- music$song.hotttnesss
```

## Logistic Regression Review
$Y \sim Binomial(n=1, p(X))$

$p(X) = \frac{exp(X\beta)}{\exp(X\beta) + 1}$

```{r}
logit <- function(x){exp(x) / (1 + exp(x))}
logistic_obj <- function(params, X, y){
  xbeta <- cbind(1, X) %*% params
  p <- logit(xbeta)
  return(-sum(dbinom(y, size=1, prob=p, log=TRUE)))
}
optim(c(1, 1), logistic_obj, X=x, y=y)
my_model <- glm(y ~ x, family=binomial(logit))
summary(my_model)
```

How to run logistic regression in R?
```{r}
train <- sample(
  c(TRUE, FALSE),
  nrow(music),
  replace=TRUE)
mod <- glm(y ~ x,
           family=binomial(link="logit"),
           subset=train)
test_probs <- predict(
  mod,
  newdata=data.frame(x=x[!train]),
  type="response")
test_pred <- test_probs > 0.5

# How good is my prediction?
same_as_data <- as.numeric(test_pred) == y[!train]
# classification error!
mean(same_as_data)
# 96% is good!
# most predictive probabilities are close to 0!

```

- $\frac{A}{A + C}$ = recall! Of all the records, that are "1", how many are our model capturing?
- $\frac{A}{A + B}$ = precision! Of all the records, that our model labels as "1", how many are truly "1"?

```{r}
alphas <- seq(0, 1, length.out=1000)
recalls <- rep(NA, length(alphas))
precisions <- rep(NA, length(alphas))
for(i in seq_along(alphas)){
  alpha <- alphas[i]
  test_pred <- test_probs > alpha
  truly1 <- y[!train] == 1
  called1 <- as.numeric(test_pred)

  recall <- sum(called1 & truly1)/ sum(truly1)
  precision <- sum(truly1 & called1) / sum(called1)
  recalls[i] <- recall
  precisions[i] <- precision
}
```


```{r}
mean(y)
pos_indices <- which(y > 0)
rep_indices <- sample(pos_indices,
                      15 * length(pos_indices), replace=TRUE)
new_y <- c(y, y[pos_indices])

starts_w_artists <- grepl("^artist", names(music))
x <- as.data.frame(scale(music[, !starts_w_artists]))
new_x <- rbind(x, x[pos_indices, ])
imbalance_mod <- glm(y ~ .,
           family=binomial(link="logit"), data=cbind(x, y))
balanced_mod <- glm(new_y ~ .,
                family=binomial(link="logit"), data=cbind(new_x, new_y))
plot(imbalance_mod$coefficients,
     balanced_mod$coefficients)
abline(a=0, b=1)
```

```{r}
step(imbalance_mod, scope=y ~ ., direction="both")
```