---
title: "Mental Health Exercise"
output: html_notebook
---

## Mental Health

Mental health in young adults are receiving more and more attention. A NYTimes opinion piece, [People in Their 20's Aren't Supposed to Be This Unhappy](https://www.nytimes.com/2023/09/27/opinion/mental-health-20s-wellness.html) claims that younger adults are experiencing a bigger decrease in happiness than other age groups.

## Task 0 - Problem identification

How would you measure unhappiness in a population?

## Taks 1 - Obtain initial dataset

[Behavior Risk Factor Surveillance System (BRFSS)](https://www.cdc.gov/brfss/index.html) includes an annual survey that tracks the number of days (out of 30) where people's "mental health was not good".


```{r}
year <- 2021
URL <- paste0("https://www.cdc.gov/brfss/annual_data/", year, "/files/LLCP", year, "ASC.zip")
URL
```

Code snippet that downloads and unzips a file (thanks OpenAI/ChatGPT 4)!

```{r}
library(httr)

resp <- httr::GET(URL)

temp_zip_path <- tempfile()
writeBin(content(resp, "raw"), temp_zip_path)

temp_dir <- tempdir()
unzipped_files <- unzip(temp_zip_path, exdir = temp_dir)
text_file_path <- unzipped_files[1]

dat <- readLines(text_file_path)

unlink(temp_zip_path)
```

```{r}
print(length(dat))
print(class(dat))
print(unique(sapply(dat, nchar)))
```

How would you describe this record of data?
```{r}
print(dat[1])
```

#### File format - Not CSV!

Each line has the same number of characters. The data is encoded into fixed length values:

See https://www.cdc.gov/brfss/annual_data/2021/llcp_varlayout_21_onecolumn.html

```{r}
layout_url <- "https://www.cdc.gov/brfss/annual_data/2021/llcp_varlayout_21_onecolumn.html"
resp <- httr::GET(layout_url)
layout_source <- httr::content(resp, "text")
```

Some quick regular expression to extract the data of interest!

```{r}
pattern <- '<td class="c data">([0-9]+)</td>\n.{0,20}MENTHLTH.{0,20}<td class="c data">([0-9]+)</td>'
dat_start <- regexpr(pattern, layout_source)
start_char <- as.numeric(sub(paste0(".*", pattern, ".*"), "\\1", layout_source))
dat_len <- as.numeric(sub(paste0(".*", pattern, ".*"), "\\2", layout_source))
stop_char <- start_char + dat_len - 1
```

```{r}
print(start_char)
print(stop_char)
```


```{r}
substr(dat[666], start_char, stop_char)
menthlth <- sapply(dat, substr, start_char, stop_char)
menthlth <- as.numeric(menthlth)
head(menthlth)
```


## Task 2 - explore initial dataset

How?

```{r}
table(menthlth)
```

### Any surprises?

Always checkout the [codebook](https://www.cdc.gov/brfss/annual_data/2019/pdf/codebook19_llcp-v2-508.HTML)!

```{r}

```


## Task 3 - Age data

- Write a function, `get_var()` using the code snippets above to get the age data.
  - Hint: with BRFSS, demographic data is starts with an underscore, e.g. "_AGE*"
  
  
```{r}
get_var <- function(var_name, layout_source){

}
```

## Task 4 - Quality check

- Check if we have the same number of values as the mental health variable
- Check if the prevalence matches those shown in the code book
  - To get the weighted percentages, you'll need `"_LLCPWT"`
  
```{r}

```

```{r}
w_loc <- get_var("_LLCPWT", layout_source)
weights <- as.numeric(sapply(dat, substr, w_loc[1], w_loc[2]))

table(age) / length(age)

print(tapply(weights, ages, sum, na.rm=TRUE) / sum(weights))
```

## Task 5 - Recreate the graph

We want age in the x-axis, % of despair within the group on the y-axis.

```{r}
despair <- menthlth == 30
summary(despair)
```

## Task 5 - Scale/clean the code above to get 2019-2022 data

We'll split into groups! Each group will:
- Work with one or two different year(s) (2015-2022)
- Enter your weighted despair percentages [here](https://docs.google.com/spreadsheets/d/1eqmO2gM-6x_P59J-aXpAe-6gvpVzq6xNYeXpeH00Hns/edit?usp=sharing)
- Ask yourself
  - Did anything change over the years in terms of handling the data?
  - How could you tell the data is "correct"
  - How can you generalize the code a bit?
  
```{r}

```

## Task 6 - How would you model the data? (Statistics - No code)

- Now you know how the data is collected, how would you model the data?
  - How will you fit the model?
- Are there alternative hypotheses to the story besides "young people are unhappy?"

## Task 7 (data science) - what will you do with this?

- What exactly will you do with the model?
- Who will you show it to
- what would you show
- what is your recommendation
- how will you show it?
