---
title: "Fisher's ANOVA"
author: "Wayne Lee"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Did Fisher Write Reproducible Papers?

For today's exercise, we'll try to reproduce one of the first papers by Fisher himself!
Fisher wanted to test the effect of fertilizer type, fertilizer amount, and fertilizer timing in a single experiment. His data can be found on the original paper on Ed under the "Resources" tab, a file called `Fisher_ANCOVA.pdf`.
We will try to reproduce his results while highlighting the importance of understanding your data types.

### Problem 0: examine the data types

Someone has entered the data from the paper for us in `/course/data/fisher_1927_straw.csv`, we will read in the data for you for now.

Please answer the following questions:
- How many rows and columns are in the dataset?
- What data type does each column have?
- Please use head() on the data to see the first 3 rows of data.
- Sanity check, does the data match the numbers from the paper?

```{r p0}
block_data <- read.csv("~/Downloads/fisher_1927_straw.csv")



```

### Problem 1: Completing the treatment information

The data is incomplete because the data frame above does not have the treatment information.
- Fertilizer amount: 0, 1, or 2 applications.
- Fertilizer type: "M" for muriate, and "S" for sulphate.
- Fertilizer timing: "E" for early and "L" for late.

We will walk you through each step below to highlight different operations and data types.

#### Problem 1.0

- Please create a vector called `f_amount` as shown in the paper (same order).
- What is the data type for this vector?

```{r p1.0}
# Hint, rep() could prove useful
rep(1:3, 2)
rep(1:3, each = 2)


```

#### Problem 1.1, two different ways to subset

A student wanted to replicate the method above to create the `f_timing` vector but noticed that the first 4 values are incorrect.
- What is the data type of the existing wrong solution?
- Decide how you want to encode the timing corresponding to "0" fertilizer then fix the vector using 2 different approaches below.
  - Method 1: use subsetting via numerical vectors (e.g. `1:2`) to replace the values corresponding to "0" fertilizer into the value you decided above.
  - Method 2: use subsetting via boolean vectors to replace the values corresponding to "0" fertilizer into the value you decided above.
    - hint: `has_no_fert <- f_amount == 0`
- Did your subsetting above change the data type of `f_timing`?

```{r}
# WRONG solution:
f_timing <- rep(c("E", "L"), 6)


```

```{r meth1}
f_timing <- rep(c("E", "L"), 6)


```

```{r meth2}
f_timing <- rep(c("E", "L"), 6)


```

#### Problem 1.1.0 Treating `[ ]` as a function

- In the example above, if `[]` was a function, in the example above, what were the inputs?
# vectors!
- Were the data types of your inputs the same across the 2 methods?
- Were the sizes of your inputs the same across the 2 methods?
  - 4
  - same length as `f_amount`
- Given your answers above:
  - Can we pass different data types to the same argument within a function?
  - If you converted `has_no_fert` to a numeric vector (e.g. `as.numeric()`), then performed the same subsetting operation, what would happen? Please explain the difference.

Please type out your answer here:


#### Problem 1.2

- Using `rep()`, please create `f_type` as shown in the paper.
- Please handle the case with "0" fertilizer the same way as you treated it above.

```{r}
x <- rep(c("S", "M"), each=2)
f_type <- rep(x, 3)
f_type[has_no_fert] <- NA
```

#### Problem 1.3

- Please add your created vectors into the original data frame using `data.frame(f_type, block_data)`, what error do you see?
- Why is this happening?

```{r}
df <- data.frame(f_amount, f_type, f_timing, block_data)
```


#### Problem 1.4 - subsetting
- Please extract the `block_totals` from `block_data` using:
  - Using subsetting 
  - Using `tail()`
- What is the data type of `block_totals` from above? Why is this the case?
- Please extract the `treatment_totals` from `block_data` using:
  - Using subsetting via a numerical vector
  - Using subsetting via a boolean vector, hint: `start_with_block <- grepl("^block", names(block_data))`
- What is the data type of `treatment_totals` from above? Why is this the case?
- Please extract the `yields` from `block_data` using subsetting (this should not have **any** aggregate values.
  - Try subsetting for rows first, then subsetting by columns by sequentially calling `[ ]`
  - Try subsetting for rows and columns with one call to `[ ]`
    - Using nuermical vectors
    - Using boolean vectors, hint: try creating a boolean from the `treatment_totals`
- Please verify that the outcomes are the same across the 3 ways of generating `yields`
```{r}
block_totals <- block_data[13, ]
block_totals <- tail(block_data, 1)
```


#### Problem 1.5 - different ways of adding variables
- Please add the treatments into the `yields` data now by
  - Trying once again, using `data.frame()` for adding `f_type`.
  - Try assigning a column that doesn't exist, e.g. `yields[, "row_ind"] <- 1:12`, for adding `f_timing`.
  - Try assigning a column that doesn't exist using `[[]]`, e.g. `yields[['row_index']] <- 1:12` for adding `f_amount`.
- Look at the ordering of the new columns
- Which of the methods above, when applied twice, will not affect the outcome? Why might this property be desirable?

```{r}

```



#### Problem 1.6 - data types is about memory footprint
- Try to replicate `f_type` 1000 times and check its memory size as a character. What if we only replicated it 10 times?
- Try to convert your replicated data above into a factor then check its memory size once again, hint `as.factor()`
- From what you observed above, are characters or factors more memory efficient?


#### Problem 2.0
- Using `yields`, please recompute the block totals and treatment totals:
  - hint: `rowSums()` and `colSums()`
- Do your values agree with the values from the paper?
  - Where did the error occur and how would you fix it?
  - Please fix the data according to your observation above.


#### Problem 2.1

In Table 3, the row that indicates "Late - Early" is the result from totalling the records that are late and subtracting the total across all early records. A similar logic applies to the other treatments as well.
Please verify if these are calculated correctly using subsetting.


#### Problems 3 Problems with Google Sheets or Excel

Please visit this [World Bank data repository](https://data.worldbank.org/indicator/NV.AGR.EMPL.KD).
- Download this data as a CSV and try to open it using the same function we opened `fisher_1927_straw.csv`. What error do you see?
- Now upload this file to Google Drive, make **NO** modifications, then re-download it as a "CSV" with a different name from your Google Drive or Excel.
- Try to read the CSV again, what happened? Is this a feature or a bug?