---
title: "joins"
author: "Wayne Lee"
date: "2/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How connected are airports?

We could rank the "connectiveness" of an airport and their associated cities by its reach around the globe. Specifically, how many cities are within 1 direct flight or within 1 transfer flight of each city.

We will perform some basic JOIN operations for this to compare the 2019 vs 2020 data.

To answer these questions, you can use the domestic and international destinations from all US airport from [this website](https://www.transtats.bts.gov/Tables.asp?DB_ID=111&DB_Name=Air%20Carrier%20Statistics%20(Form%2041%20Traffic)-%20All%20Carriers&DB_Short_Name=Air%20Carriers).
In particular, we have data from both 2019, `T_100_MARKET_DOM_INT_2020.csv`, and 2020, `T_100_MARKET_DOM_INT_2019.csv`.

### Porblem 0 data exploration

- Please verify the claim that "at least one of the origin or destination airports must be from the US".
- How many foreign countries have a direct flight with the US in 2019 (either way)?
- (No code) To calculate all possible airports within 2 flights from JFK, what information is the data missing?
- Are there countries that do not have "round trip" relationships with the US if you are restricted to only direct flights?
- Is there a one to one relationship between `ORIGIN_AIRPORT_ID` and `ORIGIN`?
  - (optional) If you find something weird, what would you say is causing it?
- How many unique airports are out there in the US?

```{r}
airports <- read.csv("~/Downloads/airports.csv")
library(tidyverse)
head(airports)
connected_to_US <- (airports$ORIGIN_COUNTRY == "US"
                    | airports$DEST_COUNTRY == "US")
all(connected_to_US)
dest_from_us_countries <- ifelse(airports$ORIGIN_COUNTRY == "US",
                            airports$DEST_COUNTRY,
                            airports$ORIGIN_COUNTRY)
is_domestic <- (airports$ORIGIN_COUNTRY == "US"
                & airports$DEST_COUNTRY == "US")
foreign_countries <- dest_from_us_countries[!is_domestic]
uniq_foreign <- unique(foreign_countries)
length(uniq_foreign)

from_us <- airports$ORIG_COUNTRY == "US"

to_us <- airports$DEST_COUNTRY == "US"
is_from_foreign <- to_us & !is_domestic
cntry_to_us <- unique(airports$ORIGIN_COUNTRY[is_from_foreign])
length(cntry_to_us)
uniq_foreign[!(uniq_foreign %in% cntry_to_us)]


head(unique(airports$ORIGIN))
head(unique(airports$ORIGIN_AIRPORT_ID))

us_airports <- unique(c(airports$DEST[to_us],
                        airports$ORIG[from_us]))
length(us_airports)
```



```{r}
user <- data.frame(
    user_id=c(1, 2, 3),
    family_name=c("Snow", "Stark", "Stark"),
    given_name=c("John", "Rob", "Arya"))
activity <- data.frame(
    user_id=c(1, 3, 1),
    page_views=c(14, 20, 1))
merge(user, activity)
```




### Problem 1

- Please remove the `*_AIRPORT_ID` variables from the data frame.
  - Do we need to deduplicate the data after removing the airport ID? Why or why not?
- At least how many airports within the US are reachable from JFK using 2 or fewer flights?
  - What about ATL? Please wrap your code from above into a function!
  - Which airport is more connected?
- How many airports are in the US?
- Sample the US airports and plot the relationship between the number of airports reachable by a direct flight vs by 2 flights (each dot is an airport). What relationship do you expect to see? (Plot it afterwards to verify!)

```{r}
ap_sans_id <- airports %>% select(-ends_with("_AIRPORT_ID")) %>% unique()

target_airport <- "ATL"
ap_sans_id %>% filter(ORIGIN == target_airport) %>% select(join_key=DEST) %>% unique() -> flight1
nrow(flight1)

airports %>% filter(to_us) %>% select(join_key=ORIGIN, DEST) %>% unique() -> flight2

connect_flights <- inner_join(flight1, flight2)
ap_reachable2 <- unique(connect_flights$DEST)

length(unique(c(ap_reachable2, flight1$join_key)))
```


### Problem 2 Impact of locking down?

- The Lombardy region in Italy was the hardest hit by COVID-19 and one of its airports is MXP.
- The Wuhan region in China is the origin of the COVID-19 pandemic and one of its airports is WUH.

(Wayne does not know if there are other airports in either region)

How would US domestic travel be impacted if we shutdown the airports with traffic **from** these two airports? This problem is intentionally open-ended. How would you quantify/visualize this?
  - Domestic travel means that your flight never leaves the US territory
  - Who is your audience and would your report change?
  
```{r}
  
```