---
title: "scraping_api"
author: "Wayne Lee"
date: "4/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scraping + Calling APIs, combining multiple data sources

You often hear that women are paid less than men for doing the same type of work. If you are a skeptic, one way to validate this is to pull some data and see for yourself!

In particular, let's look at the Fire Department in NYC and compare the payroll between men vs women in 2019.

Today's activity will be broken down into:
- Grabbing Payroll data from the [NYC OpenData](https://dev.socrata.com/foundry/data.cityofnewyork.us/k397-673e)
  - Because NYC OpenData is public, it does not require you to register for credentials before calling its API. This is definitely an exception rather than the rule. You are **highly encouraged** to go through the motions of applying for the "app token" in this document.
- Grabbing names and their associated gender from the [Sociel Security Agency](https://www.ssa.gov/oact/babynames/) under "Popular Names by Birth Year"
- A third step that will become obvious hopefully :)


### Problem 0: Getting NYC Payroll Information



- To avoid overloading Ed and NYC OpenData API, you'll need the specify several variables before calling the API. You should focus on data from the `FIRE DEPARTMENT`, in 2019, in Manhattan (you have to figure out which field this corresponds to). We will only focus on the `base_salary` and ignore other factors. You should record other variables that might help explain the variability within the data.
  - The URL is slightly hidden on the page, HINT, urls always start with `http`
  - If you're not from NYC, Queens, Bronx, Manhattan are all different Boroughs of New York City.
  - The default cap on the number of records returned is 1000, to increase this, you'll need to look for the documentation in "Paging Through Data". 
- How many people satisfy the query you sent?
- What would you explore before moving on in terms of data quality?
  - You may need to "wrangle" the data into a format that's easy to work with. During this step, you should drop the fields that are unnecessary for later analysis.
  
```{r}
library(plyr)
library(tidyverse)
library(httr)
library(xml2)


```


### Problem 1: Getting names and their associated gender

Try to get the top names for each gender for 4 different years. Again, try to divide and conquer.

- Try to load the names using your web browser, please make sure to have your "Inspect Element" or "Inspector" widget open when you're doing this.
  - You should look under the "Network" tab and look for the "Request" made out to `popularnames.cgi`, this will help you form your query automatically.
    - Notice that this is a `POST` instead of the usual `GET()`. You'll also need to pass `encode="form"` for the `body` to be formatted correctly. See `?POST` for details
- Once you get your data, you'll need to parse the HTML using `xml2::read_html`
- Write a function that takes in name and returns a recommended gender.
  - You may need to wrangle the data from above into a more friendly format
  - What happens if a name was not in the database?
  
```{r}

```

### Problem 2: Combining the 2 sources of data

- How many women are on the payroll?
- Please answer if men and women have significantly different base-salaries? What factors affect one's salary?


```{r}


```