---
title: "Violating model assumptions"
author: "Wayne Lee"
date: "3/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 0: temporally correlated errors and their impact

We know that time series is often treated differently by Statisticians. Let's learn why!

Please simualted some correlated errors according to
$$\epsilon_0 \sim N(0, 1)$$
$$\epsilon_i|\epsilon_{i-1} \sim N(\epsilon_{i-1}, 1)$$
for $i \geq 1$.

Let's set the sample size `n=50` for now.
- Please simulate 5 different sets of $\epsilon$ values according to the description above (5 * 50 values!) and 5 sets of independent $N(0, 1)$ (same length) and plot them side by side in order.
  - Please comment on the difference between the 2 sets of values.
- For the correlated errors, please visualize the distribution of $\epsilon_i$ for at least 4 different $i$ values.
  - Please compare the distributions and comment on their similarities/differences.
- Please calculate (no code) $E(\epsilon_i)$?
  - Hint: what is the relationship to $E(\epsilon_i|\epsilon_{i-1})$?
- Is $\epsilon_i$ normally distributed? Visually check this.

```{r}

```


### Question 1: What happens to regression with positively correlated errors?

Let $X = \begin{bmatrix}1 & x_1 \\ \vdots & \vdots \\ 1 & x_n \end{bmatrix}$, and $\beta = \begin{bmatrix}1 \\ 2\end{bmatrix}$. Please simulate data $Y = X\beta + \epsilon$ with a sample size $n=100$ where $\epsilon$ are the correlated erorrs.
For $x_i$, simply set these at equal intervals between 0 and 10. We will only have one observation per $x_i$ value.

- Please generate 1000 different simulations where we fit the originary least squares (OLS) model to to estimate $\beta$.
- Please comment on the $\hat{\beta}$ distribution you see across the 1000 simulations vs the distribution implied by each simulation's `summary.lm()` output.
- Please comment on the implication of this result when analyzing time series data.
- Please give an example of what $X\beta$ and $\epsilon$ could be in real life for time series. Then articulate this example with the implication above to a non-statistical but mathematically-trained colleague (e.g. physicists).

```{r}

```

### Question 2: bootstrapping for a confidence interval

How can we get a confidence interval for $\sigma^2$?

Let $Y_i \sim N(0, 1)$ for $i=1, \dots, n$ where $n=50$.
The unbiased estimate for $\hat{\sigma}^2 = \frac{1}{n-1}\sum (Y_i - \bar{Y})^2$
- Please create a bootstrap confidence interval for $\sigma^2$
- Please verify this interval, what would the alternative be if you could simulate from the distribution directly?

```{r}

```
