---
title: "confidence vs prediction intervals"
output: html_document
date: "2023-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We may also be interested in other things besides the coefficients.
- We could be interested in the true line itself, i.e. $\beta_0 + \beta_1 * X$
- We could be interested in the next measured value, i.e. $\beta_0 + \beta_1 * X + \epsilon_{new}$

Both of these would be predicted using:
$\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 * X$


But they have different uncertainties associated with them.
- For predicting the true line
  $SE(\hat{Y}) = \sigma^2 (\frac{1}{n} + \frac{(X - \bar{X})^2}{\sum_i (x_i - \bar{X})^2})$
- For predicting the next measurement
  $SE(\hat{Y}) = \sigma^2 (\frac{n + 1}{n} + \frac{(X - \bar{X})^2}{\sum_i (x_i - \bar{X})^2})$

## What does that mean?

```{r}
beta0 <- runif(1)
beta1 <- runif(1, -5, 5)

n <- 500
x <- rexp(n, 1)
y <- beta0 + beta1 * x + rnorm(n)
plot(x, y)
```

```{r}
mod <- lm(y ~ x)

pred_loc <- seq(min(x), max(x), length.out=20)
pred_y <- predict(mod, newdata = data.frame(x = pred_loc))


plot(x, y, type="n")
abline(a=beta0, b=beta1, col="blue")
points(pred_loc, pred_y, col="red")

k <- 100
new_measurements <- sapply(pred_loc, function(x) beta0 + beta1 * x + rnorm(k))
plot(x, y, type="n", ylim=range(new_measurements))
# matrix will be stacked column-wise
points(rep(pred_loc, each=k), c(new_measurements), col="#00000055", pch=16)
points(pred_loc, pred_y, col="red")
```
