---
title: "diagnostics"
output: html_document
date: "2023-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
beta0 <- -2
beta1 <- 2
sigma <- 5
n <- 100
x <- runif(n, -5, 5)
noise <- rnorm(n, sd=sigma)
```

## Standard model

```{r}
y <- beta0 + beta1 * x + noise
mod <- lm(y ~ x)

```

What were our formulas for $SE(\hat{\beta})$?

- $SE(\hat{\beta_1}) = \sqrt{\frac{\sigma^2}{\sum_i (x_i - \bar{X})^2}}$
- $SE(\hat{\beta_0}) = \sqrt{\sigma^2\left(\frac{1}{n} + \frac{\bar{X}^2}{\sum_i (x_i - \bar{X})^2}\right)}$
- $\hat{\sigma} = \sqrt{\frac{1}{n-2}\sum_i (Y_i - \hat{Y})^2} = \sqrt{\frac{1}{n-2}\sum_i (Y_i - (\hat{\beta_0} + \hat{\beta1} * x_i))^2}$

```{r}


```


## Diagnostics for Inference - not linear

```{r}
y <- beta0 + x^(beta1) + noise
mod <- lm(y ~ x)
```

## Diagnostics for Inference - noise not mean 0

```{r}
noise_odd <- rexp(n)
y <- beta0 + beta1 * x + noise_odd
mod <- lm(y ~ x)

```

## Diagnostics for inference - not independent

```{r}
noise_dep <- rnorm(1)
for(i in seq_len(n - 1)){
  noise_dep[i + 1] <- noise_dep[i] + rnorm(1, sd=0.5)
}
# sorting of x is NOT necessary for the counter example, it just makes the plots more clear
x_sorted <- sort(x)

y <- beta0 + beta1 * x_sorted + noise_dep
mod <- lm(y ~ x_sorted)

```


## Diagnostics for inference - heteroscedastic

```{r}
noise_heter <- rnorm(n, sd=sigma * abs(x / 5))
y <- beta0 + beta1 * x + noise_heter
mod <- lm(y ~ x)

```


## Prediction diagnostics - R^2

```{r}
y <- beta0 + beta1 * x + noise
mod <- lm(y ~ x)

```

## Prediction diagnostics - Cross validation

```{r}
folds <- sample(rep(1:10, 10))

```