---
title: "R Notebook"
output: html_notebook
---

A lot of this material is referenced in the [tutorial](https://leewtai.github.io/courses/stat_computing/lectures/learning_r_lln.html).

## R can replace a calculator

Try to run the code below in different ways:
- Line by line (Place your cursor at desired line then Ctrl + ENTER)
- A section (highlighting the section then Ctrl + ENTER)

```{r}
1
2*3
3 +4
3 + 4 * -1
(3 + 4) * -1
3 + 4 * - 1
```


#### Calculator functions

Functions allow us to perform complex tasks! They have
```{r}
sin(3.1415926)

sqrt(100)

log(10)

log(10, base=10)

```

## Dealing with errors

Making mistakes and learning to deal with them:
- What happens when you forget an opening parentheses?
- What happens when you forget a closing parentheses? (you can escape by triggering an error, e.g. enter 5)
- What happens when you enter non-sensible information
- What is a warning vs an error?

Components of an error message:
- 
- 

What to do when we see an error?
- 
- 

```{r, eval=FALSE}
log2)

log(2

log("hello")

mean("hello")
```


## Variable names

Variables solve the problem of:
- Similar numbers having different meanings
- Keep updates consistent 

```{r}
my_name <- "captain sparkle"
pro_noun <- "captain sparkle"

paste("Hello, my name is", my_name)
paste(pro_noun, "just said hi.")
paste("how did yuo get named", my_name)
```

## Simple data types

You need to know data types because the same function does different things
when given different data types!

```{r}
demo_logi <- TRUE
demo_num <- 123
demo_char <- "hello"

demo_logi * 2
demo_num * 2
demo_char * 2
```

## Creating data

```{r}
# Create a vector with c()
die <- c(1, 2, 3, 4, 5, 6)
# Create consecutive integers with ":"
die2 <- 1:6
# Create random variables with r???? functions
normals <- rnorm(1)
right_tail_stuff <- rexp(1)
```

## "vectorized" operations

```{r}
die * 2
die + die

# Can you explain this one?
c(1, 2, 3, 4) + c(1, 2)
```


## Functions on vectors
```{r}
mean(die)
sd(die)

```

## Subsetting and overwriting
```{r}
# oops there are two 5's!
die <- c(1, 2, 3, 4, 5, 5)

die[5] <- 6
die
```

## Nesting function calls or "the stack"

The functions always start with the innermost
```{r}
abs(die - median(die))

```

#### Exercise

How can you execute the formula below, assuming $X_1, \dots, X_6$ are the values on a die.

$$\sqrt{\frac{1}{n}\sum_i (X_i - \bar{X})^2}$$

where $\bar{X}$ is just the mean of X.
Try to use:
- vectorized operations
- the `mean()` or `sum()` function
- the `sqrt()` function

```{r}
# TYPE YOUR CODE HERE
# TYPE YOUR CODE HERE
# TYPE YOUR CODE HERE

```


## Plotting
```{r}
normals <- rnorm(100)
hist(normals)
```

## Subsetting with booleans

```{r}
normals <- rnorm(100)
is_extremes <- abs(normal) > 2
extreme_vals <- normals[is_extremes]
```

## Avoid repetition: for-loops

This happens a lot: only one value changes in the simulation!

```{r}
print("99 bottles on the wall")
print("98 bottles on the wall")
print("97 bottles on the wall")
```

Separate the changing part from the static part!

```{r}
num_bottle <- 99
paste(num_bottle, "bottles on the wall")
num_bottle <- 98
paste(num_bottle, "bottles on the wall")
num_bottle <- 97
paste(num_bottle, "bottles on the wall")
```

For-loops:

- Static code goes into the "body", i.e. between the curly brackets `{...}`
- Changing variable goes into the `for()` statement

```{r}
num_bottles <- c(99, 98, 97)
for(num_bottle in num_bottles){
  lyric <- paste(num_bottle, "bottles on the wall")
  print(lyric)
}
```

Often people like to loop over the index rather than the actual value

```{r}
num_bottles <- c(99, 98, 97)
for(i in seq_along(num_bottles)){
  print(i)
  num_bottle <- num_bottles[i]
  print(num_bottle)
  #lyric <- paste(num_bottle, "bottles on the wall")
  #print(lyric)
}
```

How to store the output from a for-loop

```{r}
lyrics <- c()
num_bottles <- c(99, 98, 97)
for(i in seq_along(num_bottles)){
  num_bottle <- num_bottles[i]
  lyric <- paste(num_bottle, "bottles on the wall")
  lyrics[i] <- lyric
}
print(lyrics)

```

