---
title: "regression"
output: html_document
date: "2023-09-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is a line?

Below is an example of the function

$$f(X) = 2 - 3 * X$$

```{r cars}
f <- function(x) {
  return(2 - 3 * x)
}
f(1)
```

- Body of the function is in "{}"
- "function" is a keyword in R to expect a function definition
- Inputs to the function are in "()"
- The variable assigned is the function name

## Vectorized operations with vectors and functions

How do we plot this?

```{r}
Xs <- -5:5
fXs <- f(Xs)
plot(Xs, fXs)
```

## What is the distance from a dot to a line?

```{r}
xy <- c(3.5, 5)
plot(Xs, fXs)
points(xy[1], xy[2], col="red")

```

## Regression's objective function

The very general "least square" objective function:
$$\sum_i (Y_i - f(X_i))^2$$
If we restrict regression to "lines" then:
$$\sum_i (Y_i - (\beta_0 + \beta_1 X_{1,i} + \beta_2 X_{2,i} + \dots))^2$$
If we restrict regression to simple regression then:
$$\sum_i (Y_i - (\beta_0 + \beta_1 X_{1,i}))^2$$
Let's try to code this last case up!
```{r}

least_squares <- function(x, y, intercept, slope){
  
  
}
```

How can we look at this?
```{r}

```

## lm() the regression function
```{r}
best_line <- lm(fXs ~ Xs)
best_line
summary(best_line)
```


## Using lm with data frames

```{r}
df <- read.csv("~/Downloads/processed_nyc_payroll_2022.csv")
best_line <- lm(regular_gross_paid ~ regular_hours, df)
summary(best_line)
```
